---
title: "Regression"
author: "Grusha Prasad"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE)
library(tidyverse)
```


## One predictor

### Generate some data 

```{r}

x = c(1:100)
y1 = x*0.5 + rnorm(100, 0, 5)
y2 = x*-2 + rnorm(100, 0, 5)
y3 = rnorm(100, 0, 5)

dat= data.frame(x=x,
                y1=y1,
                y2=y2,
                y3=y3)

```


### Fit regression model

Here is code that fits regression models for each of the three y values. What do you notice about the coefficient values and the p-values? Does this make sense? 

```{r}

fit1 = lm(y1 ~ x, data = dat)
summary(fit1)

```
```{r}

fit2 = lm(y2 ~ x, data = dat)
summary(fit2)

```

```{r}

fit3 = lm(y3 ~ x, data = dat)
summary(fit3)

```
## Two predictors

Here is code that will generate y values that are a combination of two predictors, x1 and x2. 

```{r}

x1 = c(1:100)
x2 = rep(c(1, 0), times=50)

y1 = x1*0.5 + x2*-20 + rnorm(100, 0, 5)
y2 = x*-2 + x2*0.01 + rnorm(100, 0, 10)

dat2= data.frame(x=x,
                y1=y1,
                y2=y2)

```


### Fit regression model

Here is code that fits regression models with two predictors. What do you notice about the coefficient values and the p-values? Specifically consider p-value of x2 in fit5. Does this make sense? 
```{r}

fit4 = lm(y1 ~ x1 + x2, data=dat2)
summary(fit4)

```

```{r}

fit5 = lm(y2 ~ x1 + x2, data=dat2)
summary(fit5)

```